### 2017-08-02 R/LogLikelihood_Comparison.Rmd
I went back to double check my tables of log likelihood values to make sure I hadn't entered anything incorrectly. As it turns out, I accidentally printed the physical activity models to htmlreg twice, which explains why the physical activity and mental health likelihood values were the same after adding park access into the models. I fixed the problem and added the correct values to the tables.
I removed the Northeastern census tract from Denver and also removed a handfull of census tracts from Boston that comprised mostly null values. As expected, the log likelihood values changed in Denver

---
title: "Comparison of Log Likelihood Values"
author: "Nico Boyd"
date: "August 1, 2017"
output:
  html_document: default
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width=120)
```

Greg - below I've created a table for each of our cities that compares the log likelilhood values between models. Something I found interesting was the fact that when park access is added into the models, the log likelihood values are the same for the physical activity and mental health models. Does this strike you as odd?

  > It's more odd that the log-likelihoods are the same for both model before
  the park access is added. Are you sure you didn't copy the wrong values/use the
  wrong formula?

  > *I went ahead and removed the large Northeastern census tract from Denver and also removed a handful of tracts      from Boston that comprised mostly null values. I re-ran the models and double checked myself but sure enough, the     log-likelihoods were the same for each set of models before adding park access. I discovered that the reason the log   likelihood values were the same for the physical activity and mental health models after adding park access was       that I had accidentally printed the physical activity model to htmlreg twice, so I fixed that problem and updated my   tables with the correct values. The results are less straightforward than what we saw before; see my notes below on   which models are most likely for each city and dependent variable.*

Also, quick question about model selection: since the log likelihood values are all negative, should I or should I not select the final models based on the highest **ABSOLUTE VALUE** of the log likelihood? If we're not selecting models based on absolute value, each city has two clear winners:

  > That's right; pick the least negative value.

  * Denver: SEM models 1 & 4 are most likely (both equally so) for physical activity and mental health, but least         likely for obesity; SEM models 2 & 3 are tied for most likely for obesity. The same is true of the SDM models.

  * Boston: SEM models 3 & 6 remain the most likely for all three dependent variables. However, for the SDM models, 1     & 3 are the most likely for all three variables.

  * NYC: SEM models 1 & 2 are most likely for obesity and physical activity, while all three models are equally likely     for mental health. For the SDM models, model 3 is most likely for physical activity and mental health, but Model 1     is most likely for obesity.

### Denver SEM Models
Models 1 and 4 are the most likely (both equally so) for physical activity and mental health, but least likely for obesity.
```{r denver_SEM_models, echo=FALSE}

Denver_SEM_Loglikelihoods <- data.frame(row.names=paste("Denver",1:6))
Denver_SEM_Loglikelihoods[,1] <- c(-183.33, -183.33, -183.33, -183.33, -183.33, -183.33)
Denver_SEM_Loglikelihoods[,2] <- c(-216.98, -216.98, -216.98, -216.98, -216.98, -216.98)
Denver_SEM_Loglikelihoods[,3] <- c(-134.84, -134.84, -134.84, -134.84, -134.84, -134.84)
Denver_SEM_Loglikelihoods[,4] <- c(-183.16, -183.06, -183.06, -183.16, -183.06, -183.06)
Denver_SEM_Loglikelihoods[,5] <- c(-216.86, -216.96, -216.97, -216.86, -216.96, -216.97)
Denver_SEM_Loglikelihoods[,6] <- c(-134.67, -134.84, -134.84, -134.67, -134.84, -134.84)
colnames(Denver_SEM_Loglikelihoods) <- c("Obesity_Noparks", "Physact_Noparks", "Mental_Noparks", "Obesity_Parks", "Physact_Parks", "Mental_Parks")

Denver_SEM_Loglikelihoods
```

### Denver SDM Models
Models 1 and 4 are most likely (both equally so) for physical activity and mental health, but least likely for obesity.
```{r denver_SDM_models, echo=FALSE}

Denver_SDM_Loglikelihoods <- data.frame(row.names=paste("Denver",1:6))
Denver_SDM_Loglikelihoods[,1] <- c(-158.89, -158.89, -158.89, -158.89, -158.89, -158.89)
Denver_SDM_Loglikelihoods[,2] <- c(-188.72, -188.72, -188.72, -188.72, -188.72, -188.72)
Denver_SDM_Loglikelihoods[,3] <- c(-120.84, -120.84, -120.84, -120.84, -120.84, -120.84)
Denver_SDM_Loglikelihoods[,4] <- c(-157.99, -158.54, -158.56, -157.99, -158.55, -158.56)
Denver_SDM_Loglikelihoods[,5] <- c(-185.17, -185.29, -185.34, -185.17, -185.30, -185.34)
Denver_SDM_Loglikelihoods[,6] <- c(-120.23, -120.64, -120.66, -120.23, -120.64, -120.66)
colnames(Denver_SDM_Loglikelihoods) <- c("Obesity_Noparks", "Physact_Noparks", "Mental_Noparks", "Obesity_Parks", "Physact_Parks", "Mental_Parks")

Denver_SDM_Loglikelihoods
```

### Boston SEM Models
Models 3 and 6 are most likely for all three dependent variables.
```{r boston_SEM_models, echo=FALSE}

Boston_SEM_Loglikelihoods <- data.frame(row.names=paste("Boston",1:6))
Boston_SEM_Loglikelihoods[,1] <- c(-258.16, -258.16, -258.16, -258.16, -258.16, -258.16)
Boston_SEM_Loglikelihoods[,2] <- c(-312.80, -312.80, -312.80, -312.80, -312.80, -312.80)
Boston_SEM_Loglikelihoods[,3] <- c(-211.45, -211.45, -211.45, -211.45, -211.45, -211.45)
Boston_SEM_Loglikelihoods[,4] <- c(-258.16, -258.16, -258.15, -258.16, -258.16, -258.15)
Boston_SEM_Loglikelihoods[,5] <- c(-312.80, -312.79, -312.79, -312.80, -312.79, -312.79)
Boston_SEM_Loglikelihoods[,6] <- c(-211.44, -211.43, -211.36, -211.44, -211.43, -211.36)
colnames(Boston_SEM_Loglikelihoods) <- c("Obesity_Noparks", "Physact_Noparks", "Mental_Noparks", "Obesity_Parks", "Physact_Parks", "Mental_Parks")

Boston_SEM_Loglikelihoods
```

### Boston SDM Models
Models 1 and 3 are most likely for all three dependent variables.
```{r boston_SDM_models, echo=FALSE}

Boston_SDM_Loglikelihoods <- data.frame(row.names=paste("Boston",1:6))
Boston_SDM_Loglikelihoods[,1] <- c(-230.37, -230.37, -230.37, -230.37, -230.37, -230.37)
Boston_SDM_Loglikelihoods[,2] <- c(-298.10, -298.10, -298.10, -298.10, -298.10, -298.10)
Boston_SDM_Loglikelihoods[,3] <- c(-198.01, -198.01, -198.01, -198.01, -198.01, -198.01)
Boston_SDM_Loglikelihoods[,4] <- c(-228.51, -229.42, -229.87, -228.51, -229.42, -229.87)
Boston_SDM_Loglikelihoods[,5] <- c(-297.21, -297.55, -297.73, -297.21, -297.55, -297.73)
Boston_SDM_Loglikelihoods[,6] <- c(-197.19, -197.91, -197.74, -197.20, -197.91, -197.74)
colnames(Boston_SDM_Loglikelihoods) <- c("Obesity_Noparks", "Physact_Noparks", "Mental_Noparks", "Obesity_Parks", "Physact_Parks", "Mental_Parks")

Boston_SDM_Loglikelihoods
```

### NYC SEM Models
Models 1 and 2 are most likely (both equally so) for obesity and physical activity. All three models are equally likely for mental health.
```{r nyc_SEM_models, echo=FALSE}

NYC_SEM_Loglikelihoods <- data.frame(row.names=paste("NYC   ",1:3))
NYC_SEM_Loglikelihoods[,1] <- c(-3929.85, -3929.85, -3929.85)
NYC_SEM_Loglikelihoods[,2] <- c(-4421.83, -4421.83, -4421.83)
NYC_SEM_Loglikelihoods[,3] <- c(-2716.36, -2716.36, -2716.36)
NYC_SEM_Loglikelihoods[,4] <- c(-3929.14, -3929.14, -3929.36)
NYC_SEM_Loglikelihoods[,5] <- c(-4421.80, -4421.80, -4421.82)
NYC_SEM_Loglikelihoods[,6] <- c(-2716.33, -2716.33, -2716.33)
colnames(NYC_SEM_Loglikelihoods) <- c("Obesity_Noparks", "Physact_Noparks", "Mental_Noparks", "Obesity_Parks", "Physact_Parks", "Mental_Parks")

NYC_SEM_Loglikelihoods
```

### NYC SDM Models
Model 3 is most likely for physical activity and mental health, but Model 1 is most likely for obesity.
```{r nyc_SDM_models, echo=FALSE}

NYC_SDM_Loglikelihoods <- data.frame(row.names=paste("NYC   ",1:3))
NYC_SDM_Loglikelihoods[,1] <- c(-3819.95, -3819.95, -3819.95)
NYC_SDM_Loglikelihoods[,2] <- c(-4369.45, -4369.45, -4369.45)
NYC_SDM_Loglikelihoods[,3] <- c(-2656.60, -2656.60, -2656.60)
NYC_SDM_Loglikelihoods[,4] <- c(-3808.52, -3808.53, -3812.65)
NYC_SDM_Loglikelihoods[,5] <- c(-4368.97, -4368.96, -4368.45)
NYC_SDM_Loglikelihoods[,6] <- c(-2655.48, -2655.48, -2655.36)
colnames(NYC_SDM_Loglikelihoods) <- c("Obesity_Noparks", "Physact_Noparks", "Mental_Noparks", "Obesity_Parks", "Physact_Parks", "Mental_Parks")

NYC_SDM_Loglikelihoods
```