### 2017-08-03 R/park_ls_analysis.Rmd
Ignore html.

---
title: "Park Access Analysis"
author: "Nico Boyd"
date: "August 2, 2017"
output: html_document
---

In this document I list the park_ls coefficient values for all models, indicate the presence or lack of significance, and note interesting findings that could become part of an overarching narrative in our paper.

# **Denver**

## **Denver 1** (size = .00001, distance = -5)

### Obesity Models
* OLS -> -0.05 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.05 (not significant)
* SDM -> +0.08 (not significant)  lag of park_ls = +0.19 (not significant)

### Physical Activity Models
* OLS -> +0.14 (not significant)
* SAR -> +0.14 (not significant)
* SEM -> +0.06 (not significant)
* SDM -> +0.23 (not significant)  lag of park_ls = +0.42 (**significant @ 0.01**)

  > *When my neighbors get better park access, I become less active. This model is one of the two most likely Spatial      Durbin models, the other being model 4.*

### Mental Health Models
* OLS -> +0.05 (not significant)
* SAR -> +0.05 (not significant)
* SEM -> +0.04 (not significant)
* SDM -> +0.06 (not significant)  lag of park_ls = +0.10 (not significant)

## **Denver 2** (size = .00001, distance = -15)

### Obesity Models
* OLS -> -0.02 (not significant)
* SAR -> -0.02 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> +0.01 (not significant)  lag of park_ls = +0.04 (not significant)

### Physical Activity Models
* OLS -> +0.03 (not significant)
* SAR -> +0.03 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> +0.07 (not significant)  lag of park_ls = +0.14 (**significant at 0.05**)

   > *When my neighbors get better park access, I become less active.*

### Mental Health Models
* OLS -> +0.00 (not significant)
* SAR -> +0.00 (not significant)
* SEM -> +0.00 (not significant)
* SDM -> +0.01 (not significant)  lag of park_ls = +0.02 (not significant)

## **Denver 3** (size = .00001, distance = -30)

### Obesity Models
* OLS -> -0.01 (not significant)
* SAR -> -0.01 (not significant)
* SEM -> -0.01 (not significant)
* SDM -> +0.00 (not significant)  lag of park_ls = +0.02 (not significant)

### Physical Activity Models
* OLS -> +0.01 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.00 (not significant)
* SDM -> +0.03 (not significant)  lag of park_ls = +0.07 (**significant at 0.05**)

  > *When my neighbors get better park access, I become less physically active.*

### Mental Health Models
* OLS -> +0.00 (not significant)
* SAR -> +0.00 (not significant)
* SEM -> -0.00 (not significant)
* SDM -> +0.00 (not significant)  lag of park_ls = +0.01 (not significant)

## **Denver 4** (size = .0001, distance = -5)

### Obesity Models
* OLS -> -0.05 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.05 (not significant)
* SDM -> +0.08 (not significant)  lag of park_ls = +0.19 (not significant)

### Physical Activity Models
* OLS -> +0.14 (not significant)
* SAR -> +0.14 (not significant)
* SEM -> +0.06 (not significant)
* SDM -> +0.23 (**significant at 0.05**)  lag of park_ls = +0.42 (**significant at 0.05**)

  > *When I have better park access, I become less active. When my neighbors have better park access, I become even        less active. This model is one of the two most likely Spatial Durbin models, the other being model 1.*

### Mental Health Models
* OLS -> +0.05 (not significant)
* SAR -> +0.05 (not significant)
* SEM -> +0.04 (not significant)
* SDM -> +0.06 (not significant)  lag of park_ls = +0.10 (not significant)

## **Denver 5** (size = .0001, distance = -15)

### Obesity Models
* OLS -> -0.02 (not significant)
* SAR -> -0.02 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> +0.01 (not significant)  lag of park_ls = +0.04 (not significant)

### Physical Activity Models
* OLS -> +0.03 (not significant)
* SAR -> +0.03 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> +0.07 (not significant)  lag of park_ls = +0.14 (**significant at 0.05**)

  > *When my neighbors have better park access, I become less active.*

### Mental Health Models
* OLS -> +0.00 (not significant)
* SAR -> +0.00 (not significant)
* SEM -> +0.00 (not significant)
* SDM -> +0.01 (not significant)  lag of park_ls = +0.02 (not significant)

## **Denver 6** (size = .0001, distance = -30)

### Obesity Models
* OLS -> -0.01 (not significant)
* SAR -> -0.01 (not significant)
* SEM -> -0.01 (not significant)
* SDM -> +0.00 (not significant)  lag of park_ls = +0.02 (not significant)

### Physical Activity Models
* OLS -> +0.01 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.00 (not significant)
* SDM -> +0.03 (not significant)  lag of park_ls = +0.07 (**significant at 0.05**)

  > *When my neighbors get better park access, I become less active.*

### Mental Health Models
* OLS -> +0.00 (not significant)
* SAR -> +0.00 (not significant)
* SEM -> -0.00 (not significant)
* SDM -> +0.00 (not significant)  lag of park_ls = +0.01 (not significant)

# **Boston**

## **Boston 1** (size = .00001, distance = -5)

### Obesity Models
* OLS -> +0.00 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.01 (not significant)
* SDM -> -0.30 (not significant)  lag of park_ls = -0.13 (not significant)

### Physical Activity Models
* OLS -> -0.03 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> -0.23 (not significant)  lag of park_ls = -0.26 (not significant)

### Mental Health Models
* OLS -> +0.04 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.02 (not significant)
* SDM -> -0.31 (not significant)  lag of park_ls = +0.31 (not significant)

## **Boston 2** (size = .00001, distance = -15)

### Obesity Models
* OLS -> +0.01 (not significant)
* SAR -> +0.00 (not significant)
* SEM -> +0.00 (not significant)
* SDM -> -0.08 (not significant)  lag of park_ls = -0.09 (not significant)

### Physical Activity Models
* OLS -> -0.02 (not significant)
* SAR -> -0.02 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> -0.03 (not significant)  lag of park_ls = -0.19 (not significant)

### Mental Health Models
* OLS -> +0.02 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> -0.01 (not significant)  lag of park_ls = +0.05 (not significant)

## **Boston 3** (size = .00001, distance = -30)

### Obesity Models
* OLS -> +0.01 (not significant)
* SAR -> +0.01 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> -0.03 (not significant)  lag of park_ls = -0.04 (not significant)

### Physical Activity Models
* OLS -> -0.01 (not significant)
* SAR -> -0.01 (not significant)
* SEM -> -0.01 (not significant)
* SDM -> -0.00 (not significant)  lag of park_ls = -0.09 (not significant)

### Mental Health Models
* OLS -> +0.02 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> +0.02 (not significant)  lag of park_ls = +0.03 (not significant)

## **Boston 4** (size = .0001, distance = -5)

### Obesity Models
* OLS -> +0.00 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.01 (not significant)
* SDM -> -0.30 (not significant)  lag of park_ls = -0.13 (not significant)

### Physical Activity Models
* OLS -> -0.03 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> -0.23 (not significant)  lag of park_ls = -0.26 (not significant)

### Mental Health Models
* OLS -> +0.04 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.02 (not significant)
* SDM -> -0.31 (not significant)  lag of park_ls = +0.34 (not significant)

## **Boston 5** (size = .0001, distance = -15)

### Obesity Models
* OLS -> +0.01 (not significant)
* SAR -> +0.00 (not significant)
* SEM -> +0.00 (not significant)
* SDM -> -0.08 (not significant)  lag of park_ls = -0.09 (not significant)

### Physical Activity Models
* OLS -> -0.02 (not significant)
* SAR -> -0.02 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> -0.03 (not significant)  lag of park_ls = -0.19 (not significant)

### Mental Health Models
* OLS -> +0.02 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> -0.01 (not significant)  lag of park_ls = +0.05 (not significant)

## **Boston 6** (size = .0001, distance = -30)

### Obesity Models
* OLS -> +0.01 (not significant)
* SAR -> +0.01 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> -0.03 (not significant)  lag of park_ls = -0.04 (not significant)

### Physical Activity Models
* OLS -> -0.01 (not significant)
* SAR -> -0.01 (not significant)
* SEM -> -0.01 (not significant)
* SDM -> -0.00 (not significant)  lag of park_ls = -0.09 (not significant)

### Mental Health Models
* OLS -> +0.02 (not significant)
* SAR -> +0.02 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> +0.02 (not significant)  lag of park_ls = +0.03 (not significant)

# **NYC**

## **NYC 1** (size = .00001, distance = -5)

### Obesity Models
* OLS -> -0.74 (**significant at 0.001**)
* SAR -> -0.76 (**significant at 0.001**)
* SEM -> -0.12 (not significant)
* SDM -> +0.04 (not significant)  lag of park_ls = -0.43 (**significant at 0.001**)

  > *As my neighbors' park access decreases, I become more obese. Conversely, as my neighbors get better access to         parks, I become less obese. This spatial durbin model of obesity is the most likely in NYC.*

### Physical Activity Models
* OLS -> -0.05 (not significant)
* SAR -> -0.16 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> +0.05 (not significant)  lag of park_ls = -0.15 (not significant)

### Mental Health Models
* OLS -> -0.01 (not significant)
* SAR -> -0.06 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> -0.08 (not significant)  lag of park_ls = +0.03 (not significant)

## **NYC 2** (size = .0001, distance = -5)

### Obesity Models
* OLS -> -0.74 (**significant at 0.001**)
* SAR -> -0.76 (**significant at 0.001**)
* SEM -> -0.12 (not significant)
* SDM -> +0.04 (not significant)  lag of park_ls = -0.43 (**significant at 0.01**)

  > *The OLS and SAR models indicate that I become less obese as my access to parks improves. The SDM model indicates      that as my neighbors get better access to parks, I become less obese.*

### Physical Activity Models
* OLS -> -0.05 (not significant)
* SAR -> -0.16 (not significant)
* SEM -> -0.02 (not significant)
* SDM -> +0.05 (not significant)  lag of park_ls = -0.15 (not significant)

### Mental Health Models
* OLS -> -0.01 (not significant)
* SAR -> -0.06 (not significant)
* SEM -> +0.01 (not significant)
* SDM -> -0.08 (not significant)  lag of park_ls = +0.03 (not significant)

## **NYC 3** (size = .00001, distance = -15)

### Obesity Models
* OLS -> -0.21 (**significant at 0.001**)
* SAR -> -0.21 (**significant at 0.001**)
* SEM -> -0.03 (not significant)
* SDM -> -0.03 (not significant)  lag of park_ls = -0.14 (**significant at 0.01**)

  > *The OLS and SAR models indicate that I become less obese as my access to parks improves. The SDM model indicates      that as my neighbors get better access to parks, I become less obese.*

### Physical Activity Models
* OLS -> -0.02 (not significant)
* SAR -> -0.04 (not significant)
* SEM -> -0.00 (not significant)
* SDM -> +0.01 (not significant)  lag of park_ls = -0.09 (not significant)

### Mental Health Models
* OLS -> -0.01 (not significant)
* SAR -> -0.02 (not significant)
* SEM -> -0.00 (not significant)
* SDM -> -0.02 (not significant)  lag of park_ls = -0.01 (not significant)